\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
data transmission + FFT

Traditionally, the sensor nodes are connected to computer systems with cables.
However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.

\begin{itemize}
	\item problem: viel data traffic
	\item daten können verloren gehen
	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
	\item fft
\end{itemize}

@@

\autoref{fig:UML} describes the classes of the \texttt{sensornode} package.
The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
\todo{Grafik zweitteilen - hier package sensornode}

\autoref{fig:UML} describes the classes of the \texttt{basestation} package.
The package \texttt{basestation} runs on the host computer and operates the basestation.
It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class. The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
\todo{Grafik zweitteilen - hier package basestation}

\begin{figure}[h!]
    \centering
    \includegraphics{figures/uml.pdf}
    \caption{class diagram describing the structure of the implemented SHM system}
    \label{fig:UML}
\end{figure}

>>>>

Traditionally, the sensor nodes are connected to computer systems with cables.
Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.

However, when using wireless SHM systems, special attention has to be paid to data transfer.
Immense amounts of raw data incure and have to be transmitted.
Over the air transmission is error-prone and may lead to irreparable data loss.  (klingt das zu stark? Warum ist das mit FFT nicht mehr so schlimm?)
To save transmission time and storage space, and to decrease data loss, the transmidded data can be reduced using embedded computing.
Depending on the measuring objective, first steps of data processing can be conducted decentralized directly on the sensor nodes.
Transmitting analysis results instead of raw data can reduce the amount of transmitted data exponentially.

In this paper, the natural frequency of a building structure is determined using a fast Fourier transform.
Deviations from the natural frequency indicate changes in the vibration behavior, which may be caused by structural failures.

%\begin{itemize}
%	\item problem: viel data traffic
%	\item daten können verloren gehen
%	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
%	\item fft
%\end{itemize}

@@

\autoref{fig:UML} describes the classes of the \texttt{sensor\-node} package.
The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, which are embedded directly into the sensor nodes.
The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
\todo{Grafik zweitteilen - hier package sensornode}

\autoref{fig:UML} describes the classes of the \texttt{base\-station} package.
The package \texttt{base\-station} runs on the host computer and operates the basestation.
It consists of the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
The entry point of the program is the \texttt{run()} method in the \texttt{Main\-Base} class. The \texttt{Main\-Base} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
\todo{Grafik zweitteilen - hier package basestation}

\begin{figure}[h!]
    \centering
    \includegraphics[width = \textwidth]{figures/uml.pdf}
    \caption{class diagram describing the structure of the implemented SHM system}
    \label{fig:UML}
\end{figure}
